---
title: "blblm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{blblm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Package Description
This package has functions for running little bag of bootstraps on linear regression models and logistic regression model on datasets.

Bag of little bootstrap (BLB) is a procedure that incorporates bootstrap and resampling to prodce a computationally efficient, yet robust way of estimation.

For the linear regression model (blblm), users can determine the regression coefficients (beta heads), the estimate of variance of errors (sigma), confidence interval and prediction interval on new data.
    
For logistic regression model(blblr), users can determine the regression coefficients (beta heads), as well as confidence interval on particular coefficient or prediction interval on coefficients.
    
For both function inside the package, users can also decide if they want to use parallel computing by changing the option in function.

# Linear Regression (blblm)
Download and install blblm package with the following:
```{r}
devtools::install_github("ucdavis-sta141c-2021-winter/blblm")
```

Load the package and test it out:
```{r setup}
library(blblm)
```

## Design
The primary improvement added methods that support parallelization upon users' decision.

### Parallelization
```{r, eval = FALSE}
library(furrr)
blblm <- function(formula, data, m = 10, B = 5000, parallel = FALSE, threads = 4) {
  # the function split_data put them into 
  data_list <- split_data(data, m)
  if (parallel){
    suppressWarnings(future::plan(multiprocess, workers = threads))
    estimates <- future_map(
      data_list,
      ~ lm_each_subsample(formula = formula, data = ., n = nrow(data), B = B))
  }else{
    estimates <- map(
      data_list,
      ~ lm_each_subsample(formula = formula, data = ., n = nrow(data), B = B))
  }
  res <- list(estimates = estimates, formula = formula)
  class(res) <- "blblm"
  invisible(res)
}
```

```future::plan(multiprocess, workers = threads)``` enables parallelization and in this case, the default threads/planner is at 4. if the users select parallel = TRUE. plan(multiprocess, workers = threads) is accompanied by future_map from library(furrr) for the computing

### Benchmarking

Generating data for the following examples
```{r}
devtools::load_all()
devtools::install()
set.seed(100)
n = 1000
y = rnorm(n, mean = 10, sd = 2)
x = 0.5 * y + rnorm(n, mean = 0, sd = 0.25)
z = 0.3 * y + rnorm(n, mean = 0, sd = 0.25)
data = as.data.frame(cbind(x,y,z))
```

#### Parallelization

see improvement in speed (with moderate size dataset)
```{r, echo = FALSE}
benchmark0 = bench::mark(
  # without parallelization
  blblm(y~x*z, data, m = 10, B = 5000, parallel = FALSE),
  # with parallelization
  blblm(y~x*z, data, m = 10, B = 5000, parallel = TRUE, threads = 4),
  check = FALSE)

benchmark0
```

#### 95 % Confidence Interval for Variable Coefficients

```{r, echo = FALSE}
fit_lm = blblm(y~x*z, data, m = 10, B = 5000)
fit_lm_par = blblm(y~x*z, data, m = 10, B = 5000, parallel = TRUE, threads = 4)
benchmark1 = bench::mark(
  confint(fit_lm, level = 0.95),
  confint(fit_lm, level = 0.95, parallel = TRUE, threads = 4),
  confint(fit_lm_par, level = 0.95),
  confint(fit_lm_par, level = 0.95, parallel = TRUE, threads = 4),
  check = FALSE
)
benchmark1
# ggplot2::autoplot(benchmark1)
```

#### 95% Prediction Interval

# Logistic Regression (blblr)
```{r, eval = FALSE}

```

### Benchmarking 
#### Parallelization
```{r, echo = FALSE}
benchmark3 = bench::mark(
  # without parallelization
  blblr(y~x*z, data, m = 10, B = 5000),
  # with parallelization
  blblr(y~x*z, data, m = 10, B = 5000, parallel = TRUE, threads = 4),
  check = FALSE)

benchmark3
# ggplot2::autoplot(benchmark3)
```

#### 95 % Confidence Interval for Variable Coefficients

```{r, echo = FALSE}
fit_lr = blblr(y~x*z, data, m = 10, B = 5000)
fit_lr_par = blblr(y~x*z, data, m = 10, B = 5000, parallel = TRUE, threads = 4)
benchmark4 = bench::mark(
  confint(fit_lr, level = 0.95),
  confint(fit_lr, level = 0.95, parallel = TRUE, threads = 4),
  confint(fit_lr_par, level = 0.95),
  confint(fit_lr_par, level = 0.95, parallel = TRUE, threads = 4),
  check = FALSE
)

benchmark4
#ggplot2::autoplot(benchmark4)
```

#### 95% Prediction Interval

```{r, eval = FALSE, echo = FALSE}

```